Reasoning Graphically about Quantum Computation

### 2008-03-07 papers/reasoning-graphically/paper.tex
added related work and citations.
git-svn-id: https://isaplanner.svn.sourceforge.net/svnroot/isaplanner/trunk/quantomatic@1406 f8f375fb-4f10-0410-8d47-b6fbdc7ca053

\begin{figure}[t]
$$\begin{array}{ccc}
\inlinegraphic{6em}{spider_lhs} & = & \inlinegraphic{6em}{spider_rhs}
\end{array}$$
\label{fig:informal-spider-law}\caption{An informal equation on graphs
  that expresses the Spider Theorem. }
\end{figure}

Hence any connected subgraph involving nodes of only one colour,
wherever it occurs in a graph, may be collapsed to a single vertex
labelled by a single value $\alpha$, giving a ``spider''. Informally,
this can be depicted graphically as the equation in
Figure~\ref{fig:informal-spider-law}. Conversely, a spider may be
arbitrarily divided into subspiders, provided the total in- and
out-degree is preserved, along with and the sum of the $\alpha $s.
Further, one can derive, from the spider theorem, $n$-fold versions of
many of the other equations.

@@

A crucial feature of rules like the Spider Theorem is that an arbirary
number of repetitions of a subgraph need to be matched. To support
this we introduce an operation, !-boxing (pronounced bang-boxing), on
graph representations. Given a graph representation, this introduces a
new notation by outlining a set of nodes. These nodes are said to be
in a !-box. Intuitively, the resulting !-boxed graph can be thought of
as representing a set of graphs with an arbirary number of copies of
the !-boxed nodes, where every copy connects, in the same way, to the
nodes outside the !-box. The !-boxes can also be instantiated with
zero copies. This erases all edges to and from the !-box.  The Spider
Theorem in Figure~\ref{fig:informal-spider-law} for instance is
represented with !-boxes in Figure~\ref{fig:spider-thm-patt}.

@@

Initially one might be tempted to think that the usual notion of
subgraph might surfice for expressing matching, however it allows
additional edges on nodes where, because of the tensorial semantics of
our graphs, we do not. Our work provides a restriction of the usual
subgraph definition that is sound for Compact Closed Categories.

Link Graphs and their extention to BiGraphs also encapsualte a very
different kind of semantics where they allow edges to go to multiple
nodes. An interesting area of further work would be to consider our
graph patterns for these, and other, graph based formalisms.

Graph Grammars...

Other graph rewriting...

Matrix Based Graph Tansformation...

Other Graph Transformation...

Interaction nets and interaction combinators (Lafont and others)

>>>>

\begin{figure}[t]
$$\begin{array}{ccc}
\inlinegraphic{6em}{spider_lhs} & = & \inlinegraphic{6em}{spider_rhs}
\end{array}$$

\label{fig:spider}\caption{An informal equation on graphs
  that expresses the Spider Theorem. }
\end{figure}

Hence any connected subgraph involving nodes of only one colour,
wherever it occurs in a graph, may be collapsed to a single vertex
labelled by a single value $\alpha$, giving a ``spider''. Informally,
this can be depicted graphically as the equation in
Figure~\ref{fig:spider}. Conversely, a spider may be
arbitrarily divided into subspiders, provided the total in- and
out-degree is preserved, along with and the sum of the $\alpha $s.
Further, one can derive, from the spider theorem, $n$-fold versions of
many of the other equations.

@@

A crucial feature of rules like the Spider Theorem is that an arbirary
number of repetitions of a subgraph need to be matched. To support
this we introduce an operation, !-boxing (pronounced bang-boxing), on
graph representations. Given a graph representation, this introduces a
new notation by outlining a set of nodes. These nodes are said to be
in a !-box. Intuitively, the resulting !-boxed graph can be thought of
as representing a set of graphs with an arbirary number of copies of
the !-boxed nodes, where every copy connects, in the same way, to the
nodes outside the !-box. The !-boxes can also be instantiated with
zero copies. This erases all edges to and from the !-box.  The Spider
Theorem in Figure~\ref{fig:spider} for instance is
represented with !-boxes in Figure~\ref{fig:spider-thm-patt}.

@@

There are several foundational approaches to graph transformation,
including algebraic appoaches~\cite{corradini97algebraic}, node-label
controlled~\cite{Graphgrammars83}, matrix
based~\cite{DBLP:conf/gg/VelascoL06}, and programmed graph
replacement~\cite{progress97}. These provide general ways of
interpretating graph transformations which can then be implemented to
provide machinery for a specific application.  However, systems based
on these theories do not provide machinery for the semantics of
compact closed categories. Thus their notion of matching and
replacement do not guarentee well-typed results. Furthermore, they do
not provide machinery to for rewriting of graphs with elipses
notation.

The distinctive feature of our form of graph rewriting is that the
graphs capture the structural properties of compact closed categories.
In this respect, our graphs caputure the structural properties of
compact closed categories and rewriting is compositional: it preserves
the type of the rewritten subgraph. However, the various components of
our system can be seen as instantiations for in a general graph
rewriting system: our matching provides the embedding information and
the node-data matching of the object formalism provides application
conditions and the attribute transfer function.

% Graph Grammars...
% Other graph rewriting... 
% Matrix Based Graph Tansformation...
% Other Graph Transformation...
% Interaction nets and interaction combinators (Lafont and others)
